A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-13 main.tex
[WIP] Revising Section 4.

In this section, we prove some properties of \LMD: preservation, strong normalization, confluence,
unique decomposition of evaluation contexts, and progress.
These theorem show that the type system of \LMD is strong enough to ensure a typed term of \LMD never go wrong.

@@

@@

Following three Inversion Lemmas are needed in proving later theorems.
Inversion Lemmas enable us to infer types of subterms of a term from the type of the term.
They are composed of three part.
The first part state that we can infer types of subterms.
The second and third parts state 
that when we have a type equivalence, we can infer the shape of type from the left to the right, and vice versa.
We need both of the second and the third because of the reflexivity of the type equivalence.

\begin{lemma}[Inversion Lemma for $\Pi$ type]
	If $\G \V (\lambda x:\sigma.M) : (\Pi x:\sigma'.\tau)@A$ then
	\begin{enumerate}
		\item $\G \V \sigma \E \sigma'@A$
		\item $\G ,x:\sigma@A\V M:\tau @A$
	\end{enumerate}
	\item If $\G \V \rho \E (\Pi x:\sigma.\tau) : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
	\item If $\G \V (\Pi x:\sigma.\tau) \E \rho : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$
		\item $\G \V \sigma \E \sigma' : K @A$
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$
	\end{enumerate}
\end{lemma}

@@

\begin{theorem}[Inversion Lemma for $\TW$ type]
	\begin{item}
	      \item If $\G \V \TB_\alpha M : \TW_\alpha \tau @A$ then $\G \V M : \tau @A$.
	      \item If $\G \V \rho \E  \TW_\alpha \tau : K @A$ then $\exists \tau', K, J$ such that
	      \begin{enumerate}
		      \item $\rho = \TW_\alpha \tau'$
		      \item $\G @A\V \tau \E \tau' : K @A$
	      \end{enumerate}
	      \item If $\G \V \TW_\alpha \tau \E \rho : K @A$ then $\exists \tau', K, J$ such that
	      \begin{enumerate}
		      \item $\rho = \TW_\alpha \tau'$
		      \item $\G @A\V \tau \E \tau' : K @A$
	      \end{enumerate}
	\end{item}
\end{theorem}

@@

\begin{theorem}[Inversion Lemma for $\Lambda$ type]
	\begin{item}
	      \item If $\G \V \Lambda\alpha.M : \forall\alpha.\tau @A$ then $\G \V M : \tau @A$ and $\alpha \notin \FTV(\G) \cup \FV(A)$.
	      \item If $\G \V \rho \E \forall\alpha.\tau : K @A$ then $\exists \tau', K$ such that
	      \begin{enumerate}
		      \item $\rho = \forall\alpha.\tau'$
		      \item $\G \V \tau \E \tau' : K @A$
	      \end{enumerate}
	      \item If $\G \V \forall\alpha.\tau \E \rho : K @A$ then $\exists \tau', K$ such that
	      \begin{enumerate}
		      \item $\rho = \forall\alpha.\tau'$
		      \item $\G \V \tau \E \tau' : K @A$
	      \end{enumerate}
	\end{item}
\end{theorem}

@@

Thanks to Substitution Lemmas and Inversion Lemma, we can prove Preservation easily.
Preservation ensure that any one step reduction preserves type.

\begin{theorem}[Preservation]
	If $\G\V M:\tau @A$ and $M \longrightarrow M'$, then $\G\V M':\tau @A$\\
\end{theorem}

@@

\begin{theorem}[Strong Normalization]
	If $\G\V^A M:\tau$ then there is no infinite sequence of terms $(M_i)_{i\ge1}$ and 
	$M_i \longrightarrow M_{i+1}$ for $i\ge 1$
\end{theorem}

@@

\begin{theorem}[Unique Decomposition]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau @ A$ then 1 or 2 is true.
	\begin{enumerate}
		\item $ M \in V^A$
		\item There exist an unique tuple of $B, E^A_B, R^B$ such that $M = E^A_B[R^B]$ ($B = \epsilon$ or $B = \beta$) .
	\end{enumerate}
\end{theorem}

@@

\section{Related Works}

>>>>

In this section, we prove \AI{use ``show'' because we omit proofs from the paper} basic properties of \LMD: preservation, strong normalization, confluence, and progress.
These theorem shows that the type system of \LMD is strong enough to ensure a typed term of \LMD never go wrong.
\AI{In this setting, a simple type system is already sufficient to guarantee
  ``a well-typed term never goes wrong.''  If the language has
  case analysis that uses dependent types, then
  we could say more.
}

@@

\AI{Since the order of variable declarations is significant,
  we cannot assume $z$ is always declared at the end.}

@@

The following three Inversion Lemmas are needed to prove main theorems.
As usual~\cite{TAPL}, Inversion Lemmas enable us to infer the types of subterms of a term from the type of the term.
They are composed of three parts.
The first part states that we can infer types of subterms.
The second and third parts state 
that when we have a type equivalence, we can infer the shape of type from the left to the right, and vice versa.
We need both of the second and the third because of the reflexivity of the type equivalence. \AI{Symmetry?  We can omit the third one.}

\begin{lemma}[Inversion Lemma for $\Pi$-types]
	If $\G \V (\lambda x:\sigma.M) : (\Pi x:\sigma'.\tau)@A$ then
	\begin{enumerate}
		\item $\G \V \sigma \E \sigma'@A$ and
		\item $\G ,x:\sigma@A\V M:\tau @A$.
	\end{enumerate}
	\item If $\G \V \rho \E (\Pi x:\sigma.\tau) : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$,
		\item $\G \V \sigma \E \sigma' : K @A$, and
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$.
	\end{enumerate}
	\item If $\G \V (\Pi x:\sigma.\tau) \E \rho : K @A$ then $\exists \sigma', \tau', K, J$ such that
	\begin{enumerate}
		\item $\rho = \Pi x:\sigma'.\tau'$,
		\item $\G \V \sigma \E \sigma' : K @A$, and
		\item $\G, x:\sigma@A\V \tau \E \tau' : J @A$.
	\end{enumerate}
\end{lemma}

@@

\begin{theorem}[Inversion Lemma for $\TW$-types]
	\begin{item}
	      \item If $\G \V \TB_\alpha M : \TW_\alpha \tau @A$ then $\G \V M : \tau @A$.
	      \item If $\G \V \rho \E  \TW_\alpha \tau : K @A$ then $\exists \tau', K, J$ such that
	      \begin{enumerate}
		      \item $\rho = \TW_\alpha \tau'$ and
		      \item $\G @A\V \tau \E \tau' : K @A$.
	      \end{enumerate}
	      \item If $\G \V \TW_\alpha \tau \E \rho : K @A$ then $\exists \tau', K, J$ such that
	      \begin{enumerate}
		      \item $\rho = \TW_\alpha \tau'$ and
		      \item $\G @A\V \tau \E \tau' : K @A$.
	      \end{enumerate}
	\end{item}
\end{theorem}

@@

\begin{theorem}[Inversion Lemma for $\forall$-types]
	\begin{item}
	      \item If $\G \V \Lambda\alpha.M : \forall\alpha.\tau @A$ then $\G \V M : \tau @A$ and $\alpha \notin \FTV(\G) \cup \FV(A)$.
	      \item If $\G \V \rho \E \forall\alpha.\tau : K @A$ then $\exists \tau', K$ such that
	      \begin{enumerate}
		      \item $\rho = \forall\alpha.\tau'$ and
		      \item $\G \V \tau \E \tau' : K @A$.
	      \end{enumerate}
	      \item If $\G \V \forall\alpha.\tau \E \rho : K @A$ then $\exists \tau', K$ such that
	      \begin{enumerate}
		      \item $\rho = \forall\alpha.\tau'$ and
		      \item $\G \V \tau \E \tau' : K @A$.
	      \end{enumerate}
	\end{item}
\end{theorem}

@@

Thanks to Substitution Lemmas and Inversion Lemmas, we can prove Preservation easily.
Preservations ensure that any one step reduction preserves type.

\begin{theorem}[Preservation]
	If $\G\V M:\tau @A$ and $M \longrightarrow M'$, then $\G\V M':\tau @A$.
\end{theorem}

@@

\begin{theorem}[Strong Normalization]
	If $\G\V^A M:\tau$ then there is no infinite sequence of terms $(M_i)_{i\ge1}$ and 
	$M_i \longrightarrow M_{i+1}$ for $i\ge 1$.
\end{theorem}

@@

\begin{theorem}[Unique Decomposition]
	If $x:\tau @\epsilon \notin \G$ and $\G \V M : \tau @ A$ then either
	\begin{enumerate}
		\item $ M \in V^A$, or
		\item there exist $B, E^A_B$, and $R^B$ such that $M = E^A_B[R^B]$ with $B = \epsilon$ or $B = \beta$ for some $\beta$.
	\end{enumerate}
\end{theorem}

@@

\AI{Explain the condition on $\G$.}

\section{Related Work}
